Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-06-12 main.Rmd
opening paragraph of section 4

<!-- # Why data structure for cyclic granularities exploration -->

\noindent Effective exploration and visualization require proper data structures. It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis. Moreover, one can use calendars based on application domain that define contextual system of linear and cyclic granularities.

@@

\noindent A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble] data structure. A tsibble consists of an index, key and measured variables. An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time. A linear granularity is a mapping of the index set to subsets of the time domain. For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years. A bottom granularity is represented by the index of the tsibble.

@@

\noindent All cyclic granularities can be expressed in terms of the index set, and hence, we consider the data structure in Figure \@ref(fig:data-structure) for exploration of temporal data of this kind. This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities. The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities, for example, if we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed. Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that could originate from $N_L$ periodic and aperiodic linear granularities. Any attempt to encode all or many of these cyclic granularities at the same time to develop insights on periodicity might fail or otherwise become too numerous for comprehensive human consumption. Instead, this big problem is broken down by focusing on pair of cyclic granularities ($C_i$, $C_j$) at a time for $i, j \in \{1, 2, \dots, N_C\}$. Data sets of the form <$C_i$, $C_j$, $v$> then forms the basis for exploration and analysis of the measured variable $v$.

>>>>

<!-- # Why data structure and basis for cyclic granularities exploration -->

\noindent Effective exploration and visualization benefits from well-organized data structures. @wang2019tsibble introduced a tidy data structure, tsibble, to support exploration and modeling of temporal data. This forms the basis of the structure for cyclic granularities. A tsibble consists of an index, key and measured variables. An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time. A linear granularity is a mapping of the index set to subsets of the time domain. For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years. A bottom granularity is represented by the index of the tsibble.  <!-- It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis. Moreover, one can use calendars based on application domain that define contextual system of linear and cyclic granularities.-->

@@

@@

\noindent All cyclic granularities can be expressed in terms of the index set, and hence, we introduce the data structure in Figure \@ref(fig:data-structure). This basic structure of a tsibble (index, key, measurements) is augmented by columns of cyclic granularities.\todo[inline]{think your diagram should color what's tsibble differently from what's augmented} The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities, for example, if we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed. Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that could originate from $N_L$ periodic and aperiodic linear granularities. Any attempt to encode all or many of these cyclic granularities at the same time to develop insights on periodicity might fail or otherwise become too numerous for comprehensive human consumption. Instead, this big problem is broken down by focusing on pair of cyclic granularities ($C_i$, $C_j$) at a time for $i, j \in \{1, 2, \dots, N_C\}$. Data sets of the form <$C_i$, $C_j$, $v$> then forms the basis for exploration and analysis of the measured variable $v$.
